### 2022-03-17 manuscript/figures/fourier_sampling_student.tex
TikZ figures

\begin{tikzpicture}
    \begin{semilogxaxis}[
		% Set width and height
        width = 0.54\textwidth,
        height = 0.45\textwidth,
        % Labels
        xlabel = {Degree of freedom $\alpha$},
        ylabel = {$\mathrm{MSE}~[\mathrm{dB}]$},
        y label style = {at={(0.5,1)},anchor=south,rotate=-90},
        % Grid
        grid = both,
        grid style = {thin},
        % Ticks
        ytick = {0,10,20,30,40,50,60,70},
        xtick = {20, 3, 1.5, 1},
		% Configure legend
        legend style={font=\scriptsize,at={(0.025,0.975)},anchor=north west},
  		% Reverse x plot (sparse to the right)
  		x dir = reverse,
  		x tick label style={log ticks with fixed point},
  		xmin=0.95, xmax=21,
  		ymin=3.5, ymax=75,
	]
	\foreach \column in {l2,l1,log,MMSE,NN}{
        \addplot +[fill opacity=0.5, draw opacity=0.8]
table[x=alph,y=\column]{figures/fourier_sampling_student_data.txt};
    }
    \legend{$\ell_2$,$\ell_1$,$\log$,$\mathrm{MMSE}$,CNN};
    \end{semilogxaxis}
\end{tikzpicture}